<<<<<<< HEAD
---
title: "Summary of data generation scenarios"
output: rmarkdown::pdf_document
vignette: >
  %\VignetteIndexEntry{Summary of data generation scenarios}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message = FALSE, warning = FALSE}
library(dpsurrogate)
library(splines)
library(knitr)

summary_scenario <- function(idat) {
  par(mfrow = c(2, 2))
  plot(yeff ~ seff, data = idat$rdat)
  plot(yeff ~ trtZ, data = idat$rdat)
  
  ## partial residuals
  
  fit.z <- lm(yeff ~ bs(trtZ, df = 3), data = idat$rdat)
  fit.s <- lm(yeff ~ bs(seff, df = 3), data = idat$rdat)
  
  plot(resid(fit.z) ~ idat$rdat$seff, ylab = "p resid | Z", xlab = "seff")
  plot(resid(fit.s) ~ idat$rdat$trtZ, ylab = "p resid | seff", xlab = "trtZ")
  
  summary(lm(yeff ~ seff + trtZ, data = idat$rdat))
  
}

```

We start by generating $\nu_j, Z_j, U_j$ for $j = 1, \ldots, 64$ all from independent standard normal distributions. Recall that $\nu_j$ represents the treatment effect on the potential surrogate in group $i$, and $Z_j$ is an observed treatment-level covariate, and $U_i$ is an unobserved treatment-level covariate. We consider the following scenarios for generation of $\mu_j$, the treatment effect on the clinical outcome. 

- Nonlinear (nonlinear): $\mu_j = -1 + f(\nu_j) + c_z |Z_j| + c_u U_j$, where $f$ is a linear spline basis with knots at $-1, 0, 1$ and coefficients chosen so that the trend is monotonically increasing.
- Linear (linear): $\mu_j = -1 + 1* \nu_j + c_z |Z_j| + c_u U_j$.
- Simple (simple): $\mu_j = -1 + 1* \nu_j + c_z Z_j + c_u U_j$.
- Null (null): $\mu_j = -1 +  c_z |Z_j| + c_u U$.
- Interaction (inter): $\mu_j = 1\{Z_j < 0\} * \nu_j + c_u U_j$, where $1\{\cdot\}$ is the - r function. 
- Hidden interaction (interhide): $\mu_j = 1\{U_j < 0\} * \nu_j + c_z Z_j$. 
- Surrogate value for only 1 treatment (onetrt): $\mu_j = 1\{j\in R_a\} * \nu_j + c_z Z_j + c_u U_j$ where $R_a$ is the set of indices such that the active treatment is treatment A.
- Surrogate value for 2 treatments (twotrt): $\mu_j = 1\{j\in (R_a, R_b)\} * (\nu_j - \min_j(\nu_j)) + c_z Z_j + c_u U_j$ where $R_a, R_b$ are the sets of indices such that the active treatment is treatment A and B, respectively. Additionally, the mean of $Z_j$ is changed to vary based on the treatment group (with different means in each of the 4 groups). 
- Surrogate value for multiple biomarkers (manybiom): The biomarker groups are grouped into 3 categories $B^*_j \in \{1, 2, 3\}$ of size 20, 28, 16, respectively. Then $Z_j \sim N(B^*_j - 1, 0.5^2)$ and $\mu_j = 0.25 \cdot (B^*_j - 3) \cdot (\nu_j - \min_j(\nu_j)) + c_z Z_j + c_u U_j$. In this setting, the three biomarker groups have differential surrogate value none, moderate, and strong. 

```{r inter}


settings <- rbind(data.frame(scen = c("nonlinear", "linear", "simple", 
                                      "null", "inter", "interhide",
                    "onetrt", "twotrt", "manybiom"),
           num = 99, p1 = 0, p2 = 0),
      data.frame(scen = c("nonlinear", "linear", "simple"),
                 num = 99, p1 = 0.3, p2 = 0),
      data.frame(scen = c("nonlinear", "linear", "simple"),
                 num = 99, p1 = 0.3, p2 = 0.3),
      data.frame(scen = c("inter"),
                 num = 99, p1 = 0, p2 = 0.3),
      data.frame(scen = c("interhide"),
                 num = 99, p1 = 0.3, p2 = 0))

settings$fname <- gsub(".", "", make.names(settings$scen, 
                                           unique = TRUE), fixed = TRUE)

for(i in 1:nrow(settings)){
  idat <- generate_data(effect = settings$scen[i], Zeffect = settings$p1[i], 
                        Ueffect = settings$p2[i])
  header <- sprintf("# Setting: %s, $c_z = %.1f, c_u = %.1f$ \n\n", 
              settings$scen[i], settings$p1[i], 
              settings$p2[i])
  cat(header)
  summary_scenario(idat)
  
}

```
=======
---
title: "Summary of data generation scenarios"
output: rmarkdown::pdf_document
vignette: >
  %\VignetteIndexEntry{Summary of data generation scenarios}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message = FALSE, warning = FALSE}
library(dpsurrogate)
library(splines)

summary_scenario <- function(idat) {
  par(mfrow = c(2, 2))
  plot(yeff ~ seff, data = idat$rdat)
  plot(yeff ~ trtZ, data = idat$rdat)
  
  ## partial residuals
  
  fit.z <- lm(yeff ~ bs(trtZ, df = 3), data = idat$rdat)
  fit.s <- lm(yeff ~ bs(seff, df = 3), data = idat$rdat)
  
  plot(resid(fit.z) ~ idat$rdat$seff, ylab = "p resid | Z", xlab = "seff")
  plot(resid(fit.s) ~ idat$rdat$trtZ, ylab = "p resid | seff", xlab = "trtZ")
  
  summary(lm(yeff ~ seff + trtZ, data = idat$rdat))
  
}

```

We start by generating $\nu_j, Z_j, U_j$ for $j = 1, \ldots, 64$ all from independent standard normal distributions. Recall that $\nu_j$ represents the treatment effect on the potential surrogate in group $i$, and $Z_j$ is an observed treatment-level covariate, and $U_i$ is an unobserved treatment-level covariate. We consider the following scenarios for generation of $\mu_j$, the treatment effect on the clinical outcome. 


## Nonlinear

$$\mu_j = 1 - \sum_{i = 1}^4 \beta_i B_i(\nu_j) - \gamma_1 |Z_j| - \gamma_2 U_j,$$

where $B_i(\cdot), i = 1, \ldots, 4$ is a linear B-spline basis with knots at $-1, 0, 1$, $\beta = (0, .6, 3, 5)$. 

When $\gamma_1 = \gamma_2 = 0$, we have the following relationship:

```{r nonlin00}
idat <- generate_data(effect = "nonlinear", Zeffect = FALSE, Ueffect = FALSE)
summary_scenario(idat)
```


When $\gamma_1 = 0.25, \gamma_2 = 0$, we have the following relationship:

```{r nonlin10}
idat <- generate_data(effect = "nonlinear", Zeffect = TRUE, Ueffect = FALSE)
summary_scenario(idat)
```


When $\gamma_1 = 0, \gamma_2 = 0.25$, we have the following relationship:

```{r nonlin01}
idat <- generate_data(effect = "nonlinear", Zeffect = FALSE, Ueffect = TRUE)
summary_scenario(idat)
```



When $\gamma_1 = \gamma_2 = 0.25$, we have the following relationship:

```{r nonlin11}
idat <- generate_data(effect = "nonlinear", Zeffect = TRUE, Ueffect = TRUE)
summary_scenario(idat)
```


## Linear


$$\mu_j = 1 - 1 \nu_j - \gamma_1 |Z_j| - \gamma_2 U_j.$$


When $\gamma_1 = \gamma_2 = 0$, we have the following relationship:

```{r lin00}
idat <- generate_data(effect = "linear", Zeffect = FALSE, Ueffect = FALSE)
summary_scenario(idat)
```


When $\gamma_1 = 0.25, \gamma_2 = 0$, we have the following relationship:

```{r lin10}
idat <- generate_data(effect = "linear", Zeffect = TRUE, Ueffect = FALSE)
summary_scenario(idat)
```


When $\gamma_1 = 0, \gamma_2 = 0.25$, we have the following relationship:

```{r lin01}
idat <- generate_data(effect = "linear", Zeffect = FALSE, Ueffect = TRUE)
summary_scenario(idat)
```



When $\gamma_1 = \gamma_2 = 0.25$, we have the following relationship:

```{r lin11}
idat <- generate_data(effect = "linear", Zeffect = TRUE, Ueffect = TRUE)
summary_scenario(idat)
```

## Null


$$\mu_j = 0 - 0 \nu_j - \gamma_1 |Z_j| - \gamma_2 U_j.$$

When $\gamma_1 = \gamma_2 = 0$, we have the following relationship:

```{r null00}
idat <- generate_data(effect = "null", Zeffect = FALSE, Ueffect = FALSE)
summary_scenario(idat)
```


When $\gamma_1 = 0.25, \gamma_2 = 0$, we have the following relationship:

```{r null10}
idat <- generate_data(effect = "null", Zeffect = TRUE, Ueffect = FALSE)
summary_scenario(idat)
```


When $\gamma_1 = 0, \gamma_2 = 0.25$, we have the following relationship:

```{r null01}
idat <- generate_data(effect = "null", Zeffect = FALSE, Ueffect = TRUE)
summary_scenario(idat)
```



When $\gamma_1 = \gamma_2 = 0.25$, we have the following relationship:

```{r null11}
idat <- generate_data(effect = "null", Zeffect = TRUE, Ueffect = TRUE)
summary_scenario(idat)
```
>>>>>>> new computer
